---
title: "Google Analytics Application Shiny"
output: html_document
runtime: shiny
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/R Files")
```

## Introduction

The purpose of this exercise is to interactively query the Google Analytics API. 

Shiny Applications are written using two scripts that reference one another. These two scripts are **user interface (UI)** and **server scripts**. Though separate, these scripts are kept in the same folder and contain complementary information. 

First, install and access the Google Analytics package in R. 

```{r, eval=FALSE}
install.packages("googleAnalyticsR")
```
```{r}
library(googleAnalyticsR)
```

Next, read in and name the data. For this example, we are using sample data from a website.  

```{r}
GAdf <- read.csv("GoogleAnalyticsdf") 
```


## The UI Script

First, it is necessary to install and access the rgdal package. This package allows us to select a **date** and **time range** and view a resulting text summary or plot as an output.

The following is the ui script for the Google Analytics Application:

```{r}
# Always remember to access the shiny package.
library(shiny)

# The following section is concerned with inputs. 
shinyUI(fluidPage(
  titlePanel("Google Analytics"),
  sidebarLayout(
    sidebarPanel(
      dateRangeInput(inputId = "dateRange", label = "Date range", 
                     start = "2013-05-01"),

checkboxGroupInput(inputId = "domainShow",
                         label = "Show NHS and other domain
                                (defaults to all)?",
                         choices = list("NHS users" = "nhs.uk",
                                        "Other" = "Other"),
                         selected = c("nhs.uk", "Other")),
      hr(),

radioButtons(inputId = "outputRequired", label = "Output required", 
                   choices = list("Average session" = "meanSession", 
                                  "Users" = "users", "Sessions" = "sessions")),

checkboxInput("smooth", label = "Add smoother?", value = FALSE)),

# The second section of the ui script lays out the output panel.
mainPanel(
      tabsetPanel( 
        tabPanel("Summary", textOutput("textDisplay")),
        tabPanel("Trend", plotOutput("trend")),
        tabPanel("Map", plotOutput("ggplotMap"))
      )
    )
  )
))
```

Note the meaning of the above functions/elements:

Function/Element         | Use
------------------------ | --------------------------------------------------------------------
**titlePanel()**         | gives the application a title   
**sidebarLayout()**      | builds a simple setup (controls on left; output on right)
**sidebarPanel()**       | configures sidebar layout
**dateRangeInput()**     | provides start- and end-date widgets 
**checkboxGroupInput()** | allows user to select from a number of tick boxes  
**radioButtons()**       | provides radio buttons, allowing for the selection of only one option
**checkboxInput()**      | gives a tick box that reads TRUE when selected and FALSE when unselected 
**mainPanel()**          | configures the main panel section
**tabsetPanel()**        | sets up a tabbed output 

(Note: leaving the end-date blank in the dateRangeInput() function will default to today's date, as shown in this example.)  

## The Server Script

Several packages are needed to complete the server.R script. These include:

Function/Element         | Use
------------------------ | --------------------------------------------------------------------
**dplyr**                | filter() command; chained instructions
**ggplot2**              | build ggplot objects
**rgdal**                | selection of date and time ranges and output of text or plot
**RColorBrewer**         | pcolor palettes for thematic maps 

```{r, eval=FALSE}
# Install the dplyr, ggplot2, rgdal and RColorBrewer packages. 
install.packages("dplyr")
install.packages("ggplot2")
install.packages("rgdal")
install.packages("RColorBrewer")
```
```{r}
# Access the dplyr, ggplot2, rgdal, shiny and RColorBrewer packages. 
library(dplyr)
library(ggplot2)
library(rgdal)
library(shiny)
library(RColorBrewer)

# Build the Reactive Objects.

shinyServer(function(input, output) {
  
  load("gadf.Rdata")

  passData <- reactive({
    
    firstData <- filter(gadf, date >= input$dateRange[1] & date <= input$dateRange[2])
    
    if(!is.null(input$domainShow)){
      
      firstData <- filter(firstData, networkDomain %in% input$domainShow)
      
    }
    
    return(firstData)
    
  })
 
  if(!is.null(input$domainShow)){
      
      firstData <- filter(firstData, networkDomain %in% input$domainShow)
      
  }

  return(firstData)
    
  })

# Build the Outputs 

## The first tab of output is the Text Output (text summary of the data).

output$textDisplay <- renderText({
    paste(
      length(seq.Date(input$dateRange[1], input$dateRange[2], by = "days")),
      " days are summarised. There were", sum(passData()$users),
      "users in this time period."
    )
  })

## The next tab of output contains Trend Graphs (specifies trend graph in specified varaible).

output$trend <- renderPlot({
    
    groupByDate <- group_by(passData(), YearMonth, networkDomain) %>%
      summarise(meanSession = mean(sessionDuration, na.rm = TRUE), 
                users = sum(users),
                newUsers = sum(newUsers), sessions = sum(sessions))
    
    groupByDate$Date <- as.Date(paste0(groupByDate$YearMonth, "01"), format = "%Y%m%d")
    
    thePlot <- ggplot(groupByDate, 
                      aes_string(x = "Date", y = input$outputRequired, 
                                 group = "networkDomain", colour = "networkDomain")) +
      geom_line()

    if(input$smooth){
      
      thePlot <- thePlot + geom_smooth()
    }
    
    print(thePlot)
    
  })

## The final tab of output: Choropleth Map 

output$ggplotMap <- renderPlot ({
    
    groupCountry <- group_by(passData(), country)
    
    groupByCountry <- summarise(groupCountry, meanSession = mean(sessionDuration), 
                                users = log(sum(users)), sessions = log(sum(sessions)))

### Merge data with objects to link the map and variables of interest 
    
    world <- readOGR(dsn=".", layer="world_country_admin_boundary_shapefile_with_fips_codes")
    
    countries <- world@data
    
    countries <- cbind(id = rownames(countries), countries)
    
    countries <- merge(countries, groupByCountry, by.x = "CNTRY_NAME", by.y = "country", all.x = TRUE)
    
    map.df <- fortify(world)
    
    map.df <- merge(map.df, countries, by = "id")
    
    ggplot(map.df, aes(x = long, y = lat, group = group)) +
      geom_polygon(aes_string(fill = input$outputRequired)) +
      geom_path(colour = "grey50") +
      scale_fill_gradientn(colours = rev(brewer.pal(9, "Spectral")),
                           na.value = "white") +
      coord_fixed() + labs(x = "", y = "")
    
  })
```

Note the meanings of the above functions/elements:

Function/Element                        | Use
--------------------------------------  | ----------------------------------------------------------------------------
**load()**                              | where data preparation takes place
**reactive({...})**                     | marks whatever contained inside it as reactive; generates a **reactive object** 
**filter(gadf ...)**                    | filters data by the dates that the user is interested in using 
**if(!is.null(...))**                   | returns NULL when a group of checkboxes has no return values; ensures that only checkboxes that are selected are included in the filter 
**{filter(firstData ...)}**             | filters data using specified, within-function network domains 
**return(firstData)**                   | returns a value from a function and closes the function
**length()**                            | finds the length between specified vector of dates as specified in the ui.R script
**seq.Date(...)**                       | generates regular sequence(s) of dates
**output$trend <- renderPlot({...})**   | defines output as a **reactive plot**
**groupbyDate <- group_by()**           | groups by variables 
**summarise()**                         | produces summarized session durations, users and new users, grouped according to the                                              | desired groupings (in this case, month/year and network domain groupings)
**groupByDate <- as.date()**            | creates a date from the YearMonth variable; adds "01" to end of variable; converts to R Date format
**thePlot <- ggplot()**                 | main setup of plot 
**aes_string()**                        | allows passage of string(s) from variable(s) straight into ggplot (in this case, from input$outputsRequired)
**if(input$smooth){...}**               | tests value of smoothing checkbox and adds a smoothing line if TRUE value is returned 
**print()**                             | prints the plot 
**output$ggplotMap <- renderPlot({})**  | sets up a *reactive plot**
**log()**                               | logs variables; compresses the scale and enhances readabiliy 
**world <- readOGR(...)**               | carries out processing of map 
**ggplot()**                            | builds a ggplot object
