---
title: Intro Shiny Example
author: 
output: html_document
runtime: shiny
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/R Files")
```

## Building an Interactive Histogram

### Step 1. Open a new R Script. Save this R Script as a .Rmd file. 

### Step 2. Format YAML Header. 

The following six lines should precede the rest of the text in the RMarkdown document. (In other words, copy and paste the first six lines of this document.)

Be sure to specify the document's runtime as shiny, as this signals R to download the necessary packages. 

### Step 3. Knit the document. 

To knit the document, press Ctrl + Shift + K. Knitting the document will cause another window to open. This window will contain the title and author information as provided in the YAML Header. Close this window. 

### Step 4. Enter code chunk(s).

Following is an interactive, reactive histogram: 

``` {r, echo = FALSE}
sliderInput("sampleSize", label = "Size of sample", 
            min = 10, max = 100, value = 50, step = 1)
renderPlot({hist(runif(input$sampleSize))})
```


* **sliderInput():** (input code) Defines an input widget and allows users to select from a range of values.   
* **renderPlot():** (output code) Produces a reactive plot by utilizing the functions within it.   
* **hist():** Builds a histogram.  
* **runif():** produces random numbers, given a quantity argument. (In this case, the argument is input$sampleSize.) 

### Step 5. Knit the document.

To knit the document, press Ctrl + Shift + K. Again, knitting the document will prompt a separate window to appear. This window will contain the interactive histogram.